---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# distiller

<!-- badges: start -->
[![R-CMD-check](https://github.com/asenetcky/distiller/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/asenetcky/distiller/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/asenetcky/distiller/graph/badge.svg)](https://app.codecov.io/gh/asenetcky/distiller)
<!-- badges: end -->

## Disclaimer
The author of this package is in no away affiliated with the CDC.
This package was created without their input or approval and is only meant
to soften some of the sharp edges of the EPHT submission process.  Users will
still need to wrangle their data into the format this package requires, and
will still need to implement much of the CDC logic in their wrangle.  This
package also does _not_ submit your data to the CDC's portal.  It _only_
creates the XML document that the CDC's tooling requires.  There will be _no_
guarantees that this package will work for your specific use-case, and if
their are changes to the submission process or required formats this package
may not be updated to reflect those changes.  Use at your own risk.


## Motivation
As a newbie who has to submit data to the CDC's EPHT program, I was
dismayed to find out that the documentation is buried under many layers 
inside their SharePoint.  It is also highly fragmented, convoluted and in many
cases, conflicts with itself.  

My goal is to make this process easier and reproducible for myself, and others.

So who is this highly specific package for?

  *  Do you submit data to the CDC's EPHT program?
  *  Do you use R? Or are interested in incorporating R into your workflow?
  *  Do you struggle with the CDC's EPHT documenation and/or tooling?
  *  Do you want to make your submission process more reproducible?
  
  If you answered yes to the first question and any of the others, 
  then this package might be for you.
  

## Installation

You can install the development version of distiller from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("asenetcky/distiller")
```

## Example

Here is a basic example of how to use it:

```{r example}
library(distiller)

# Take you already-wrangled data
# note the specific variable names
data <-
    mtcars |>
    dplyr::rename(
      month = mpg,
      agegroup = cyl,
      county = disp,
      ethnicity = hp,
      health_outcome_id = drat,
      monthly_count = wt,
      race = qsec,
      sex = vs,
      year = am
    ) |>
    dplyr::select(-c(gear, carb))

# And your metadata
  content_group_id <- "AS-HOSP"
  mcn <- "1234-1234-1234-1234-1234"
  jurisdiction_code <- "two_letter_code"
  state_fips_code <- "1234"
  submitter_email <- "submitter@email.com"
  submitter_name <- "Submitter Name"
  submitter_title <- "Submitter Title"

# And then make your xml document
  make_xml_document(
    data,
    content_group_id,
    mcn,
    jurisdiction_code,
    state_fips_code,
    submitter_email,
    submitter_name,
    submitter_title,
    check_first = TRUE #optional sanity check on your data and metadata
    )
```
