---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# distiller

<!-- badges: start -->
[![R-CMD-check](https://github.com/asenetcky/distiller/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/asenetcky/distiller/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/asenetcky/distiller/graph/badge.svg)](https://app.codecov.io/gh/asenetcky/distiller)
<!-- badges: end -->

Right off the bat I want to say I am in no way affiliated with the
CDC. As someone who now has to submit data to the CDC's EPHT program I was
dismayed to find out that the documentation is highly convoluted and in many
cases, conflicts with itself.  When reaching out to support, they tell you
to read the documentation, which is not helpful.  Also, some of the provided
tooling _does not_ work if you go through the instructions provided and do
everything in the order as presented.  Basically folks who have been doing
this for a while are fine, or have made their peace with this process and then
all the newbies, like me are just endlessly sending data to the test portal
and waiting and waiting and waiting for results. 

My goal is to make this process easier and reproducible for myself, and others.

So who is this highly specific package for?

  *  Do you submit data to the CDC's EPHT program?
  *  Do you hate documentation that shows variables in the examples that don't 
  exist in the data dictionary or vice-verse?
  *  Do you hate it when the how-to-guides very clearly show the variables 
  in UPPERCASE but then the portal kicks your data back because it expects it
  to be in lowercase?
  *  Are you baffled by the XML design decisions and lack of consistency in the
  formatting between the different types of facilities?
  *  Do you hate it when you follow the directions verbatim when using CDC
  tooling, but it constantly throws null-pointer exceptions at you because
  the very first step you were supposed to do was at the end of the instruction
  guide?  
  *  Do you hate it even more when you reach out to support they tell you to 
  read the manual, instead of fixing the manual or even their own bug?
  
  If you answered yes to any of these questions, then this package might be for you
  

## Installation

You can install the development version of distiller from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("asenetcky/distiller")
```

## Example

For the six or so people this package applies to, here is a basic example of how to use it:

```{r example}
library(distiller)


# Take you already-wrangled data
# note the specific variable names
data <-
    mtcars |>
    dplyr::rename(
      month = mpg,
      agegroup = cyl,
      county = disp,
      ethnicity = hp,
      health_outcome_id = drat,
      monthly_count = wt,
      race = qsec,
      sex = vs,
      year = am
    ) |>
    dplyr::select(-c(gear, carb))

# And your metadata
  content_group_id <- "AS-HOSP"
  mcn <- "1234-1234-1234-1234-1234"
  jurisdiction_code <- "two_letter_code"
  state_fips_code <- "1234"
  submitter_email <- "submitter@email.com"
  submitter_name <- "Submitter Name"
  submitter_title <- "Submitter Title"

# And then make your xml document
  make_xml_document(
    data,
    content_group_id,
    mcn,
    jurisdiction_code,
    state_fips_code,
    submitter_email,
    submitter_name,
    submitter_title,
    check_first = TRUE #optional sanity check on your data and metadata
    )
```
